
```{r warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, fig.height = 3)
```

```{r include=FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(maptools)
library(knitr)
df <- read_csv("~/Documents/My Consultancies/IDLG District Governor's Report/DG data (March 19) v5.csv")
pop <- read_csv("~/Documents/My Consultancies/IDLG District Governor's Report/Afghan_opium_survey_2017_cult_CLEAN.csv")
names(pop) <- c("province", "dist", "year", "cultivation")
pop$year <- pop$year-621
```

```{r warning=FALSE, include=FALSE}
df$education <- factor(df$education, 
                       levels = c("Master's degree", "Bachelor degree", "14th grade", "School certificate", "School incomplete", "Private education"))
df$majEthnic <- ifelse(df$ethnic == "Pashtun", "Pashtun", 
                       ifelse(df$ethnic == "Tajik", "Tajik", ifelse(df$ethnic == "Hazara", "Hazara", ifelse(df$ethnic == "Uzbek", "Uzbek", ifelse(!is.na(df$ethnic), "Others", NA)))))
df$majEthnic <- factor(df$majEthnic, levels = c("Pashtun", "Tajik", "Hazara", "Uzbek", "Others"))
df$native <- as.character(df$native)
df$native <- ifelse(df$native == "TRUE", "Native", ifelse(df$native == "FALSE", "Non-native", NA))
df$competitiveHiring <- ifelse(df$dghiring=="Competitive", "Competitive", ifelse(is.na(df$dghiring), NA, "Uncompetitive")) %>% factor(ordered = T)
df <- mutate(df, startyear = year(df$startdate), startAge = startyear - birthyear, 
             ageGrps = ifelse(startAge<=30, 1, ifelse(startAge<=40, 2, ifelse(startAge<=50, 3, ifelse(startAge<=60, 4, 5)))))
df$ageGrps <- factor(df$ageGrps, labels = c("Up to 30", "31-40", "41-50", "51-60", "More than 60"))
df$secure <- factor(df$secure, labels = c("Collapsed","Insecure","Somewhat Insecure","Somewhat Secure","Secure"), 
                    levels = c("Not Active","Insecure","Somewhat Insecure","Somewhat Secure","Secure"), ordered = T)
df$native <- df$pob == df$province
df <- mutate(df, border=factor(border, labels=c("Not Frontier District", "Frontier District")), 
             ports=factor(ports, labels=c("Without Port", "With Port")), 
             mines=factor(mines, labels=c("Without Mine", "With Mine")), 
             grade=factor(grade, labels=c("Grade 1", "Grade 2", "Grade 3")), 
             suburb=factor(suburb, labels=c("Not Peri-Urban Districts", "Peri-Urban Districts")), 
             highway=factor(highway, labels=c("Without Highway", "With Highway")), 
             native=factor(native, labels=c("Non-Native", "Native")),
             secureBinary=case_when(secure %in% c("Insecure", "Somewhat Insecure") ~ "Insecure", 
                                    secure %in% c("Somewhat Secure", "Secure") ~ "Secure", 
                                    secure=="Collapsed" ~ "Collapsed"))
df$secureBinary <- factor(df$secureBinary, levels = c("Secure", "Insecure", "Collapsed"), ordered = T)

df2 <- data.table(select(df, -interval))
df2 <- df2[, list(prov,district,firstname,fathername,pob,birthyear,ageGrps,majEthnic,competitiveHiring,education,native,polpartyaffiliation,pastpolpartyaffiliation,province,dist,status,secure,secureBinary,security,border,ports,mines,grade,suburb,highway, date = seq(startdate, enddate, by = "day")), by = 1:nrow(df2)]
df2$age <- as.numeric((as.Date(df2$date) - as.Date(paste0(df2$birthyear, "-07-01")))/365)
df2$ageGrps2 <- if_else(df2$age<30.5, 1, if_else(df2$age<40.5, 2, if_else(df2$age<50.5, 3, if_else(df2$age<60.5, 4, 5))))
df2$ageGrps2 <- factor(df2$ageGrps2, labels = c("Up to 30", "31-40", "41-50", "51-60", "More than 60"))
df2$year <- year(df2$date)
df2 <- full_join(df2, pop, c("province", "dist", "year"))

nugIncept <- as.Date("1393-07-07")
incept <- format.Date(nugIncept, "%d/%m/%Y") %>% as.character()
nugHired <- filter(df, startdate > nugIncept & !is.na(firstname))
shp <- readShapePoly("~/Desktop/ShinyApps/Province mapmaker/AFG_adm1.shp")
```



## DISTRICT GOVERNORS

### Age of District Governors

```{r include=FALSE}
ageNug <- filter(df2, date %in% c(nugIncept, as.Date("1396-12-30"))) %>% .[, c("date", "age")] %>% 
  group_by(date) %>% summarize(age = mean(age, na.rm = T)) %>% mutate(age = format(round(age, 1), nsmall = 1))
ageGps <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$ageGrps, useNA = "ifany")} %>% prop.table() %>% data.frame
under41 <- ageGps %>% slice(1:2) %>% select(Freq) %>% sum() %>% {format(round(.*100, 1), 1)}
bw41n60 <- ageGps %>% slice(3:4) %>% select(Freq) %>% sum() %>% {format(round(.*100, 1), 1)}
```

Although there is not a clear-cut stance on age of district governors, there is a lot of interest in knowing whether younger governors perform better, or older and more experienced governors are more suitable. In this section we explore the trends, and correlations between age and other important indicators related to districts and district governors. In Part 2 of this report, using rigorous methods of inferential statistics, we attempt at answering the questions about performance. On average, there is a small increase in the average age of district governors since inception of NUG, from `r ageNug[1,2]` to `r ageNug[2,2]` years. In terms of age groups, `r under41`% of governors are up to 40 years old and `r bw41n60`% are between 41 and 60 years old.

```{r echo=FALSE, fig.cap="Age of district governors across time"}
df2[, c("prov", "district", "date", "ageGrps2")] %>% group_by(date, ageGrps2) %>% tally %>% 
  ggplot(aes(date, n, fill = ageGrps2)) + geom_area() + theme_bw() + 
  geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Age")
```

At the end of 1396, the oldest district governor was Haji Hamisha Gul Shinwari from Dur Baba district of Nangarhar with 69 years of age. The youngest district governor was Mohammad Khan in Fersi district in Herat province with 23 years of age. Below table lists the oldest and youngest district governors at the end of 1396.


```{r echo=FALSE}
filter(df, enddate == as.Date("1396-12-30")) %>% arrange(birthyear) %>% 
  mutate(age = 1397-birthyear) %>% select(province, dist, age) %>% head(5) %>% 
  kable(col.names = c("Province", "District", "Age"), caption = "Oldest District Governors")

filter(df, enddate == as.Date("1396-12-30")) %>% arrange(desc(birthyear)) %>% 
  mutate(age = 1397-birthyear) %>% select(province, dist, age) %>% head(5) %>% 
  kable(col.names = c("Province", "District", "Age"), caption = "Youngest District Governors")
```

The age of district governors appears to have relationship with some districts’ geo-political characteristics. On average, more strategic districts which are classified as grade 1 and districts that are adjacent to the provincial centers have had older district governors in the past 4 years. Districts with natural resources, ports, or frontier districts along Pakistan border have younger governors on average. Younger governors are also associated with greater opium poppy cultivation. Districts’ security condition or having highway cross the district does not appear to have clear relationship with the age of district governors. The Figure below shows the relationship between district security status and age of their governors at the national level, across time.

```{r echo=FALSE, fig.cap="Relationship between district governors' age and districts' security status"}
df2[, c("province", "district", "date", "ageGrps2", "secureBinary")] %>% 
  group_by(date, ageGrps2, secureBinary) %>% tally %>% filter(!is.na(secureBinary)) %>%
  ggplot(aes(date, n, fill = ageGrps2)) + geom_area(position = "fill") + facet_wrap(~secureBinary) + 
  theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") +
  labs(x="", y="Ratio", fill="Age", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r include=FALSE}
ethAge <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$majEthnic, .$ageGrps, useNA = "ifany")} %>% 
  prop.table(1) %>% data.frame()
hazaraOld <- ethAge %>% filter(Var1 == "Hazara") %>% slice(4:5) %>% select(Freq) %>% sum() %>% {format(round(.*100, 1), 1)}
tajikYng <- ethAge %>% filter(Var1 == "Tajik") %>% slice(1:3) %>% select(Freq) %>% sum() %>% {format(round(.*100, 1), 1)}
```

In terms of ethnicity, district governors from Hazara ethnic groups are considerably older, with nearly half (`r hazaraOld`%) of them older than 50 years of age. Conversely, Tajik district governors have on average younger, with two-third (`r tajikYng`%) of them are 50 years old or younger. Additionally, there is a strong relationship between competitive appointments and younger age, and higher educational attainment, among district governors with more younger governors being recruited in competitive process.

```{r echo=FALSE, fig.cap="Relationship between district governors' age and appointment type"}
df2[, c("province", "district", "date", "ageGrps2", "competitiveHiring")] %>% 
  group_by(date, ageGrps2, competitiveHiring) %>% tally %>% filter(!is.na(competitiveHiring)) %>%
  ggplot(aes(date, n, fill = ageGrps2)) + geom_area(position = "fill") + facet_wrap(~competitiveHiring) + 
  theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") +
  labs(x="", y="Ratio", fill="Age", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```


### Educational Attainment of District Governors

```{r include=FALSE}
UnivEdu <- filter(df2, date %in% c(as.Date("1393-01-01"), as.Date("1396-12-30"))) %>% 
  {table(.$date, .$education, useNA = "ifany")} %>% prop.table(1) %>% {.*100} %>% data.frame() %>% 
  filter(Var2 %in% c("Master's degree", "Bachelor degree")) %>% group_by(Var1) %>% summarise(sum(Freq)) %>%
  {round(.[[2]], 1)}
ppInc <- (UnivEdu[2]-UnivEdu[1])
pInc <- ((UnivEdu[2]-UnivEdu[1])/UnivEdu[1]) %>% {round(.*100, 1)}
```

There has been a steady increase in educational level of district governors in the last four years. For instance, in the beginning of the 1393, `r UnivEdu[1]`% of all district governors held a university degree such as bachelor's or master's degrees. At the end of 1396, this figure was `r UnivEdu[2]`%, which indicates a `r pInc`% (or `r ppInc` percentage points) increase in district governors with a university degree.

```{r echo=FALSE, fig.cap="Educational attainment of district governors across time"}
df2[, c("prov", "district", "date", "education")] %>% group_by(date, education) %>% tally %>% 
  ggplot(aes(date, n, fill = education)) + geom_area() + theme_bw() + 
  geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Education")
```

```{r echo=FALSE, fig.cap="Change in educational level of district governors since inception of NUG"}
filter(df2, date %in% c(nugIncept, as.Date("1396-12-30"))) %>% 
  .[, c("education", "date")] %>% table(useNA = "ifany") %>% prop.table(2) %>% data.frame %>%
  ggplot(aes(education, Freq, fill = date)) + geom_col(position = "dodge") + theme_bw() + 
  labs(x="", y="Ratio", fill="") + theme(axis.text.x = element_text(angle = 90, vjust=.5, hjust=1))
```

```{r include=FALSE}
UnivEduGrd1 <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$grade, .$education)} %>% prop.table(1) %>% {.*100} %>% 
  data.frame %>% filter(Var2 %in% c("Master's degree", "Bachelor degree"), Var1 == "Grade 1") %>% {sum(.$Freq)} %>% round(1)
UnivEduPeriUrban <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$suburb, .$education)} %>% prop.table(1) %>% {.*100} %>%
  data.frame %>% filter(Var2 %in% c("Master's degree", "Bachelor degree")) %>% 
  group_by(Var1) %>% summarize(university = round(sum(Freq),1))
UnivEduFrontier <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$border, .$education)} %>% prop.table(1) %>% {.*100} %>%
  data.frame %>% filter(Var2 %in% c("Master's degree", "Bachelor degree")) %>%
  group_by(Var1) %>% summarize(university = format(round(sum(Freq), 1), nsmall = 1))
```

While there is overall improvement in educational attainment, the improvement is more pronounced among governors of grade 1 districts (`r UnivEduGrd1`% with university degree). Furthermore, there has been significant educational gains among governors of not peri-urban districts (`r UnivEduPeriUrban[1,2]`% with university degree) that almost equals peri-urban districts (`r UnivEduPeriUrban[2,2]`% with university degree), and frontier districts (`r UnivEduFrontier[2,2]`% with university degree) surpassing not frontier districts (`r UnivEduFrontier[1,2]`% with university degree). 

```{r echo=FALSE, fig.cap="Relationship between district governors' education and districts' grade"}
df2[, c("province", "district", "date", "education", "grade")] %>% group_by(date, education, grade) %>% tally %>% 
  filter(!is.na(grade)) %>% ggplot(aes(date, n, fill = education)) + geom_area(position = "fill") + theme_bw() + 
  geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Education", title="") + 
  facet_wrap(~grade) + theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Relationship between district governors' education and peri-urban districts"}
df2[, c("province", "district", "date", "education", "suburb")] %>% group_by(date, education, suburb) %>% tally %>% 
  filter(!is.na(suburb)) %>% ggplot(aes(date, n, fill = education)) + geom_area(position = "fill") + facet_wrap(~suburb) + 
  theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Education", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Relationship between district governors' education and frontier districts"}
df2[, c("province", "district", "date", "education", "border")] %>% group_by(date, education, border) %>% tally %>% 
  filter(!is.na(border)) %>% ggplot(aes(date, n, fill = education)) + geom_area(position = "fill") + theme_bw() + 
  facet_wrap(~border) + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Education", title="") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r include=FALSE}
UnivEduHiring <- df2[df2$date == as.Date("1396-12-30"), ] %>% {table(.$education, .$competitiveHiring, useNA = "ifany")} %>% 
  prop.table(2) %>% {.*100} %>% data.frame() %>% filter(Var1 %in% c("Master's degree", "Bachelor degree")) %>% 
  group_by(Var2) %>% summarise(round(sum(Freq), 1))
UnivEduNative <- df2[df2$date == as.Date("1396-12-30"),] %>% {table(.$education, .$native, useNA = "ifany")} %>%
  prop.table(2) %>% {.*100} %>% data.frame() %>% filter(Var1 %in% c("Master's degree", "Bachelor degree")) %>%
  group_by(Var2) %>% summarise(Freq = format(round(sum(Freq), 1), nsmall = 1))
UnivEduEthnic <- df2[df2$date == as.Date("1396-12-30"),] %>% {table(.$education, .$majEthnic, useNA = "ifany")} %>%
  prop.table(2) %>% {.*100} %>% data.frame() %>% filter(Var1 %in% c("Master's degree", "Bachelor degree")) %>%
  group_by(Var2) %>% summarise(university = round(sum(Freq), 1)) %>% arrange(desc(university)) %>%
  filter(Var2 != "Others")
```

There is a negative relationship between educational attainment and age of district governors, with on average younger district governors having higher educational attainment. Furthermore, majority (`r UnivEduHiring[1,2]`%) of competitively recruited district governors have at least a university degree, compared to `r UnivEduHiring[2,2]` percent of uncompetitively recruited district governors. Non-native district governors are also more likely to be university educated (`r UnivEduNative[1,2]`%) than native district governors (`r UnivEduNative[2,2]`%). In terms of ethnic groups, most (`r UnivEduEthnic[1,2]`%) of governors from `r UnivEduEthnic[1,1]` ethnic group have at least a university degree. Conversely, only `r UnivEduEthnic[4,2]` percent of governors from `r UnivEduEthnic[4,1]` ethnic group have a university degree.

```{r echo=FALSE, fig.cap="Relationship between district governors' education and their ethnicity"}
df2[, c("province", "district", "date", "education", "majEthnic")] %>% group_by(date, education, majEthnic) %>% tally %>% 
  filter(!is.na(majEthnic)) %>% ggplot(aes(date, n, fill = education)) + geom_area(position = "fill") + facet_wrap(~majEthnic) + 
  theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Education", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

### Ethnicity of district governors

```{r include=FALSE}
ethnicity96 <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$majEthnic, useNA = "ifany")} %>% 
  prop.table() %>% {round(.*100, 1)} %>% data.frame() %>% arrange(desc(Freq)) %>% 
  mutate(Freq = format(Freq, nsmall = 1))
majethnic96 <- ethnicity96 %>% filter(Var1 %in% c("Pashtun", "Tajik", "Uzbek", "Hazara")) %>% 
  mutate(Freq = as.numeric(Freq)) %>% {sum(.[[2]])}
```

In the last four years, district governors belonged to 14 different ethnic groups. Majority (`r majethnic96`%) of them belong to 4 major ethnic groups. `r ethnicity96[1,1]` ethnic governors accounts for the largest share of district governors (`r ethnicity96[1,2]`%), followed by `r ethnicity96[2,1]` (`r ethnicity96[2,2]`%), `r ethnicity96[3,1]` (`r ethnicity96[3,2]`%) and `r ethnicity96[4,1]` (`r ethnicity96[4,2]`%). Comparing to when the National Unity Government was established, the overall share of ethnic groups has not changed significantly.

```{r echo=FALSE, fig.cap="Change in ethnicity of District governors since inception of NUG"}
filter(df2, date %in% c(nugIncept, as.Date("1396-12-30"))) %>% group_by(date, majEthnic) %>% tally() %>%
  ggplot(aes(majEthnic, n, fill = as.factor(date))) + geom_col(position = "dodge") + theme_bw() +
  labs(x="", y="Number of District Governors", fill="")
```

```{r include=FALSE}
ethSecurity <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$majEthnic, .$secureBinary, useNA = "ifany")} %>% 
  prop.table(1) %>% data.frame() %>% filter(Var2 %in% c("Insecure", "Collapsed"), !is.na(Var1), Var1 != "Others") %>% 
  group_by(Var1) %>% summarise(insecurity = round(sum(Freq)*100,1)) %>%
  arrange(insecurity)
ethFront <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$majEthnic, .$border, useNA = "ifany")} %>% 
  prop.table(2) %>% data.frame() %>% filter(!is.na(Var1), Var2 == "Frontier District", Var1 != "Others") %>% 
  arrange(desc(Freq)) %>% mutate(Freq = format(round(Freq*100, 1), 1))
ethHW <- filter(df2, date == as.Date("1396-12-30")) %>% {table(.$majEthnic, .$highway, useNA = "ifany")} %>% 
  prop.table(2) %>% data.frame() %>% filter(!is.na(Var1), Var2 == "With Highway", Var1 != "Others") %>% 
  arrange(desc(Freq)) %>% mutate(Freq = format(round(Freq*100, 1), 1))
```

Nearly half (`r ethSecurity[4,2]`%) of districts which are governed by ethnic `r ethSecurity[4,1]` governors are insecure (or collapsed). The insecurity is much lower in districts governed by ethnic `r ethSecurity[3,1]` (`r ethSecurity[3,2]`%), `r ethSecurity[2,1]` (`r ethSecurity[2,2]`%), and `r ethSecurity[1,1]` (`r ethSecurity[1,2]`%) governors. It is important to note that there is no evidence for a causal relationship, and that security has relationship with other factors (see section 1.1.3). Furthermore, majority (`r ethFront[1,3]`%) of governors in frontier districts belong to ethnic `r ethFront[1,1]` governors, followed by `r ethFront[2,1]` (`r ethFront[2,3]`%) and `r ethFront[3,1]` (`r ethFront[3,3]`%) governors, and no ethnic Hazara governor serves in frontier districts. Furthermore, in terms of highway access, districts with ethnic `r ethHW[1,1]` governors have `r ethHW[1,3]`% access, followed by `r ethHW[2,1]` (`r ethHW[2,3]`%), `r ethHW[3,1]` (`r ethHW[3,3]`%), and `r ethHW[4,1]` (`r ethHW[4,3]`%).

```{r echo=FALSE, fig.cap="Relationship between district governors' ethnicity and districts' security"}
df2[df2$date == as.Date("1396-12-30"),] %>% {table(.$majEthnic, .$secureBinary, useNA = "ifany")} %>% 
  prop.table(1) %>% {round(.*100, 1)} %>% data.frame() %>% mutate(Var2 = fct_rev(Var2)) %>% 
  filter(!is.na(Var1)) %>% ggplot(aes(Var1, Freq, fill = Var2)) + geom_col() + 
  scale_fill_manual(values = c("black", "firebrick2", "darkolivegreen4"), na.value = "grey") +
  theme_bw() + labs(x="", y="Ratio", fill="Security")
```

```{r echo=FALSE, fig.cap="Relationship between district governors' ethnicity and districts' access to highway"}
df2[, c("province", "district", "date", "majEthnic", "highway")] %>% group_by(date, majEthnic, highway) %>% 
  tally %>% filter(!is.na(highway)) %>% ggplot(aes(date, n, fill = majEthnic)) + geom_area(position = "fill") + 
  facet_wrap(~highway) + theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") + 
  labs(x="", y="Ratio", fill="Age", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Relationship between district governors' ethnicity and districts' having a border with a neighboring country"}
df2[, c("province", "district", "date", "majEthnic", "border")] %>% group_by(date, majEthnic, border) %>% 
  tally %>% filter(!is.na(border)) %>% ggplot(aes(date, n, fill = majEthnic)) + geom_area(position = "fill") + 
  facet_wrap(~border) + theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") + 
  labs(x="", y="Ratio", fill="Ethnicity", title="") + theme(axis.text.x = element_text(angle = 90, vjust = .5))
```

```{r include=FALSE}
nativity <- filter(df2, date == as.Date("1396-12-30")) %>% filter(!is.na(firstname)) %>%
  count(native) %>% mutate(perc = format(round((n/sum(n))*100, 1), nsmall = 1))
ethAge <- filter(df2, date == as.Date("1396-12-30")) %>% filter(!is.na(majEthnic)) %>% group_by(majEthnic) %>% 
  summarise(age = format(round(mean(age, na.rm = T), 1), nsmall = 1)) %>% arrange(desc(age))
```

While only a fraction (`r nativity[1,3]`%) of district governors are non-native, ethnic Pashtun governors have relatively greater representation. In terms of age, ethnic `r ethAge[1,1]` district governors have the highest average age (`r ethAge[1,2]` years). In terms of type of recruitment, a growing share of ethnic Tajik governors are recruited through uncompetitive processes in the past two years, while only a negligible portion of ethnic Hazara district governors are recruited through an uncompetitive process.

```{r echo=FALSE, fig.cap="Relationship between Nativity and Ethnicity"}
df2[, c("province", "district", "date", "majEthnic", "native")] %>% count(date, majEthnic, native) %>% filter(!is.na(native)) %>%
  group_by(date, native) %>%  mutate(perc = n / sum(n)) %>% arrange(date, native) %>% ggplot(aes(date, perc, fill = majEthnic)) +
  geom_area() + facet_wrap(~native) + theme_bw() + geom_vline(xintercept = nugIncept, col = "navy")
```

```{r echo=FALSE, fig.cap="Average age of district governors by ethnicity"}
df2[, c("province", "district", "date", "majEthnic", "age")] %>% group_by(date, majEthnic) %>% summarise(age = mean(age, na.rm = T)) %>%
  filter(!is.na(majEthnic)) %>% ggplot(aes(date, age, col = majEthnic)) + geom_line() + theme_bw() + 
  geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Avearge Age", col="Ethnicity")
```

```{r echo=FALSE, fig.cap="Relationship between district governors' ethnicity and type of their appointment / selection"}
df2[, c("province", "district", "date", "majEthnic", "competitiveHiring")] %>% group_by(date, majEthnic, competitiveHiring) %>% tally %>% 
  filter(!is.na(competitiveHiring)) %>% ggplot(aes(date, n, fill = majEthnic)) + geom_area(position = "fill") + 
  facet_wrap(~competitiveHiring) + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Ethnicity", title="") + 
  theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

### Nativity of district governors

Approximately 80% of districts have governors who are native to the province, this has not changed much in the past 4 year.

```{r echo=FALSE, fig.cap="Ethnicity of District governors across time"}
df2[, c("prov", "district", "date", "native")] %>% group_by(date, native) %>% tally %>% 
  ggplot(aes(date, n, fill = native)) + geom_area() + theme_bw() + 
  geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="Age")
```

Native district governors tend to have more presence in the peri-urban districts. Another notable difference is districts bordering Pakistan, which have more non-native governors, while other districts bordering countries other than Pakistan have more native governors than the average. The share of native governors is also higher among districts that are classified as highly insecure. Moreover, almost all governors of districts with ports and districts classified as grade 1 district used to be native until mid-1395. Districts with highways are also more likely to have native governors. The relationship between security and whether district governor is native does not appear to be strong, as shown by the figure below.

```{r echo=FALSE, fig.cap="Relationship between district governors' nativity and districts' security status"}
df2[, c("province", "district", "date", "native", "secure")] %>% group_by(date, native, secure) %>% tally %>% 
  filter(!is.na(secure))   %>% ggplot(aes(date, n, fill = native)) + geom_area(position = "fill") + theme_bw() + 
  facet_wrap(~secure) + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r include=FALSE}
nativeAge <- filter(df2, date == as.Date("1396-12-30")) %>% group_by(native) %>% summarise(age = mean(age, na.rm = T)) %>%
  mutate(age = format(round(age, 1), 1))
```

Native district governors are on average older (`r nativeAge[2,2]` years) than non-native district governors (`r nativeAge[1,2]` years). Non-native governors, being on average younger, also have a higher level of education on average, and are appointed more likely to be appointed via a competitive process.

```{r echo=FALSE, fig.cap="Relationship between nativity and appointment type of district governors"}
df2[, c("province", "district", "date", "native", "competitiveHiring")] %>% 
  group_by(date, native, competitiveHiring) %>% tally %>% filter(!is.na(native)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + facet_wrap(~native) + 
  theme_bw() + geom_vline(xintercept = nugIncept, col = "navy") + labs(x="", y="Ratio", fill="", title="") + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```


### Appointment / Selection of District Governors

```{r include=FALSE}
nugHired <- df %>% filter(!is.na(firstname), startdate > nugIncept)
nAppoint <- nugHired %>% {.[, c("province", "district", "firstname", "startdate")]} %>% mutate(startdate = year(startdate)) %>% group_by(province, startdate) %>% tally %>% 
  spread(startdate, n) %>% mutate(appmt = sum(`1393`, `1394`, `1395`, `1396`, na.rm = T)) %>% arrange(desc(appmt))
recTot <- sum(nAppoint[[6]])
hired <- table(nugHired$competitiveHiring, useNA = "ifany") %>% prop.table() %>% {round(.*100, 1)} %>% data.frame()
dismiss <- nugHired %>% filter(enddate < as.Date("1396-12-30")) %>% nrow()
dismissLen <- nugHired %>% filter(enddate < as.Date("1396-12-30")) %>% mutate(int = as.numeric(enddate - startdate)) %>% select(int) %>% {sum(.)/dismiss} %>% round(1)
```

Since the inception of the National Unity Government (NUG) on `r incept` till the end of 1396, 199 district governors were appointed in all provinces either through competitive or uncompetitive procedures. However, due to reasons unexplored in this report, district governors have been reshuffled by the central government or the provincial governors. In total, the reshufflings icnrease the number of appointments to `r recTot` since the inception of the NUG. This report uses the _"true"_ district governors, the ones that have been serving as district governors irrespective of whether they were appointed formally or reshuffled, for the analysis. The table below provides the number of appointments by province across time since `r incept`.

```{r echo=FALSE}
options(knitr.kable.NA = "")
nAppoint %>% kable(col.names = c("Province", "1393*", "1394", "1395", "1396", "Total Appointments"), 
                   caption = "Number of Appointments of District Governors Since the Inception of NUG on 1393/07/07")
```

The following graph provides a visual presentation of number of new district governors appointments across time. The red line indicates the inception of the NUG.

```{r echo=FALSE, fig.cap="Appointments / selection of new district governors across time"}
seq <- seq.Date(as.Date("1393-01-01"), as.Date("1396-12-30"), by = "day")
rec <- numeric()
for (i in seq) {
  rec <- c(rec, filter(df, !is.na(firstname), startdate == as.Date(i, origin = "1970-01-01")) %>% nrow())
}
data.frame(seq, rec) %>% filter(seq > as.Date("1393-01-01")) %>% ggplot(aes(seq, rec)) + geom_line(alpha = .5) + geom_point(size = .4, alpha = .4) + geom_vline(xintercept = nugIncept, col = "red") + theme_bw() +
  labs(x = "", y = "Number of Appointments")
```

Since the inception of NUG, in total `r recTot` district governors were appointed either as acting governors (`r hired[2,2]`%) or selected through competitive process (`r hired[1,2]`%)--remaining `r hired[3,2]` percent is unclear. Of those governors or acting governors, `r dismiss` district governors or acting governors were replaced by another by the end of 1396. On average, the length of their service was close to a year (`r dismissLen` days), with some as short as one month and as lengthy as more than 2 years. The following graph illustrates the frequency of length of service among `r dismiss` district governors who started under the NUG and were laid off.

```{r echo=FALSE, fig.cap="Duration of Appointment for district governors that were appointed since the NUG and let go before end of 1396"}
nugHired %>% filter(enddate < as.Date("1396-12-30")) %>% mutate(int = enddate - startdate) %>% arrange(desc(int)) %>% select(int) %>%
  ggplot(aes(int)) + geom_histogram(fill = "navy") + theme_bw() + labs(x = "Duration in Days", Y = "Count")
```

Overall, the number of district governors appointed / selected via competitive process seem to be on decline since in last year, despite an initial increased in competitive recruitments after formation of NUG.

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors across time"}
df2[, c("province", "dist", "date", "competitiveHiring")] %>% group_by(date, competitiveHiring) %>% tally %>% ggplot(aes(date, n, fill = competitiveHiring)) + geom_area() + theme_bw() +
  labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy")
```

The following graphs, the relationship between the type of appointments of district governments and district characteristics are explored using graphs. Note, the vertical line indicates the inception date of the NUG.

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and security status"}
df2[, c("province", "dist", "date", "competitiveHiring", "secure")] %>% group_by(date, competitiveHiring, secure) %>% tally %>% filter(!is.na(secure)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~secure) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and districts grade"}
df2[, c("province", "dist", "date", "competitiveHiring", "grade")] %>% group_by(date, competitiveHiring, grade) %>% tally %>% filter(!is.na(grade)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~grade) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and district's access to highway"}
df2[, c("province", "dist", "date", "competitiveHiring", "highway")] %>% group_by(date, competitiveHiring, highway) %>% tally %>% filter(!is.na(highway)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~highway) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and districts' being adjacent to a provincial center"}
df2[, c("province", "dist", "date", "competitiveHiring", "suburb")] %>% group_by(date, competitiveHiring, suburb) %>% tally %>% filter(!is.na(suburb)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~suburb) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and districts' having a border with a neighboring country"}
df2[, c("province", "dist", "date", "competitiveHiring", "border")] %>% group_by(date, competitiveHiring, border) %>% tally %>% filter(!is.na(border)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~border) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and districts' having mineral deposits / mines"}
df2[, c("province", "dist", "date", "competitiveHiring", "mines")] %>% group_by(date, competitiveHiring, mines) %>% tally %>% filter(!is.na(mines)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~mines) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

Furthermore, the relationship between the types of appointments of district governors and their characteristics such as education, age, ethnicity and nativity are explored using graphs. The vertical line in graphs indicates the inception date of the NUG.

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and their age"}
df2[, c("province", "dist", "date", "competitiveHiring", "ageGrps2")] %>% group_by(date, competitiveHiring, ageGrps2) %>% tally %>% filter(!is.na(ageGrps2)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~ageGrps2) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and their education"}
df2[, c("province", "dist", "date", "competitiveHiring", "education")] %>% group_by(date, competitiveHiring, education) %>% tally %>% filter(!is.na(education)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~education) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and their ethnicity"}
df2[, c("province", "dist", "date", "competitiveHiring", "majEthnic")] %>% group_by(date, competitiveHiring, majEthnic) %>% tally %>% filter(!is.na(majEthnic)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~majEthnic) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```

```{r echo=FALSE, fig.cap="Share of competitive and uncompetitive district governors and their nativity"}
df2[, c("province", "dist", "date", "competitiveHiring", "native")] %>% group_by(date, competitiveHiring, native) %>% tally %>% filter(!is.na(native)) %>%
  ggplot(aes(date, n, fill = competitiveHiring)) + geom_area(position = "fill") + theme_bw() + facet_wrap(~native) + labs(x = "", y = "Count", fill = "Type of\nAppointment") + geom_vline(xintercept = nugIncept, col = "navy") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5), plot.title = element_text(hjust = .5))
```




